---
title: "DiabeticPreprocessing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing dataset

We import the dataset and see which variables have missing values
```{r}
#Set working directory to the one with the data in the following way: setwd("YOUR/PATH")  //NOTICE ORIENTATION OF BAR
diabetic_data <- read.csv("diabetic_data.csv", na.strings = c("?"))
names(which(sapply(diabetic_data, anyNA)))
```

First we want to see the % of missings that each variable has

```{r echo=FALSE}
print(paste0("race: ",(sum(is.na(diabetic_data$race))/dim(diabetic_data)[1])*100))
print(paste0("weight: ",(sum(is.na(diabetic_data$weight))/dim(diabetic_data)[1])*100))
print(paste0("payer_code: ",(sum(is.na(diabetic_data$payer_code))/dim(diabetic_data)[1])*100))
print(paste0("medical_specialty: ",(sum(is.na(diabetic_data$medical_specialty))/dim(diabetic_data)[1])*100))
print(paste0("diag_1: ",(sum(is.na(diabetic_data$diag_1))/dim(diabetic_data)[1])*100))
print(paste0("diag_2: ",(sum(is.na(diabetic_data$diag_2))/dim(diabetic_data)[1])*100))
print(paste0("diag_3: ",(sum(is.na(diabetic_data$diag_3))/dim(diabetic_data)[1])*100))
```

The first decision is to work only with individuals whose weight has been recorded. We notice that these NAs might not be random but structural, but nevertheless we choose this scope of work for our study, as a new criteria over our data: "Each individual should have their weight recorded."
```{r}
diabetic_data <- diabetic_data[!is.na(diabetic_data$weight), ]
n<-dim(diabetic_data)[1]
```

Our dataset has been significantly reduced, and now we have `r n` observations. We re-evaluate the missing values once again.
```{r echo=FALSE}
names(which(sapply(diabetic_data, anyNA)))
print(paste0("race: ",(sum(is.na(diabetic_data$race))/dim(diabetic_data)[1])*100))
print(paste0("payer_code: ",(sum(is.na(diabetic_data$payer_code))/dim(diabetic_data)[1])*100))
print(paste0("medical_specialty: ",(sum(is.na(diabetic_data$medical_specialty))/dim(diabetic_data)[1])*100))
print(paste0("diag_2: ",(sum(is.na(diabetic_data$diag_2))/dim(diabetic_data)[1])*100))
print(paste0("diag_3: ",(sum(is.na(diabetic_data$diag_3))/dim(diabetic_data)[1])*100))
```

Our next step is to input missing values and, since those variables with NAs are all categorical, we will create a new factor named <var>_unknown for each one of them, substituting the missing values by it. This will help for computation and visualization procedures.

```{r}
diabetic_data$race <- factor(diabetic_data$race, levels= c(levels(diabetic_data$race),"race_unknown"))
diabetic_data$race[is.na(diabetic_data$race)] <- "race_unknown"

diabetic_data$payer_code <- factor(diabetic_data$payer_code, levels= c(levels(diabetic_data$payer_code),"payer_code_unknown"))
diabetic_data$payer_code[is.na(diabetic_data$payer_code)] <- "payer_code_unknown"

diabetic_data$medical_specialty <- factor(diabetic_data$medical_specialty, levels= c(levels(diabetic_data$medical_specialty),"specialty_unknown"))
diabetic_data$medical_specialty[is.na(diabetic_data$medical_specialty)] <- "specialty_unknown"

diabetic_data$diag_2 <- factor(diabetic_data$diag_2, levels= c(levels(diabetic_data$diag_2),"diag_2_unknown"))
diabetic_data$diag_2[is.na(diabetic_data$diag_2)] <- "diag_2_unknown"

diabetic_data$diag_3 <- factor(diabetic_data$diag_3, levels= c(levels(diabetic_data$diag_3),"diag_3_unknown"))
diabetic_data$diag_3[is.na(diabetic_data$diag_3)] <- "diag_3_unknown"
```

Now we have no NAs (or at least, they have been reskinned). We want to see the levels of our factors:

```{r}
summary(diabetic_data)
```

We notice that some variables have not been imported correctly as factors(admission_source_id, admission_type_id, discharge_disposition_id). We fix it.
```{r}
diabetic_data$admission_source_id = as.factor(diabetic_data$admission_source_id)
diabetic_data$admission_type_id = as.factor(diabetic_data$admission_type_id)
diabetic_data$discharge_disposition_id = as.factor(diabetic_data$discharge_disposition_id)
```

We notice several things from this last summary. Firstly, we have too many medicines, most of them apparently uninformative for the subset of data we are dealing with. We also have too many factors over diagnostics, and after a look at the meaning of those factors we notice they are not even ordinal and would not be too useful in their current form (too disperse and the factors' semantical relationship is not expressed).

From these observations, we decide that the next steps for modelling our data are the following:
  - Collapse the drugs variables, droping most of the unsignificant ones. Our criteria has been to maintain insuline and metformine as they are (since they are both important for diabetes and well distributed over the factors) and to group all of the rest into a new variable 'other_meds', which will be binary, with value 'no_more_meds' if all of the other drugs have 'No' value and 'takes_more_meds' otherwise.
  - Reduce factors over diagnostics, grouping them semantically. That is, instead of using specific diagnostics, we will map them to the general area they are related to, such as circulatory, respiratory etc. For that, we already have a table made by experts of the field.
  - Rename the variables and factors if they are too long, for the sake of having a cleaner look.

We will first work on collapsing the drugs. To recapitulate, we have decided to name this new variable 'other_meds', which will be a binary variable with factors:'takes_more_meds' and 'no_more_meds' (which will help distinguish it from other qualitative vars when we do the PCA).

```{r}
diabetic_data$other_meds <- as.factor(with(diabetic_data, ifelse(repaglinide=='No' & nateglinide == 'No'& chlorpropamide == 'No'& glimepiride == 'No'& acetohexamide == 'No'& glipizide == 'No'& glyburide == 'No'& tolbutamide == 'No'& pioglitazone == 'No'& rosiglitazone == 'No'& acarbose == 'No'& miglitol == 'No'& troglitazone == 'No'& tolazamide == 'No'& examide == 'No' & citoglipton == 'No', 'no_more_meds', 'takes_more_meds')))

table(diabetic_data$other_meds)

```

As we see above, this new variable is quite well balanced (which implies it will probably be informative as well). By reducing the number of variables (which were imbalanced between their factors), we hope to reduce the execution time of our algorithms while keeping a (now not-so-informative as before) record over the rest of medicines which appear to be more punctual than metformin and insuline.
Having checked that, we now drop the compiled medicine variables from the dataset. We'll also drop the variable max_glu_serum, since we have observed it has no descriptive capabilities (all values are 'None').

```{r echo=FALSE}
cat("Previous data:")
colnames(diabetic_data)
diabetic_data_1 <- diabetic_data[, -which(names(diabetic_data) %in% c("repaglinide","nateglinide","chlorpropamide","glimepiride","acetohexamide","glipizide","glyburide","tolbutamide","pioglitazone","rosiglitazone","acarbose","miglitol","troglitazone","tolazamide","examide","citoglipton","glyburide.metformin","glipizide.metformin","glimepiride.pioglitazone","metformin.rosiglitazone","metformin.pioglitazone", "max_glu_serum"))]

cat("\nNew data:\n")
colnames(diabetic_data_1)
diabetic_data <- diabetic_data_1
```

We will now reduce the factors, using the table provided by experts. We first look over the factor distribution of diag_1
```{r}
plot(diabetic_data$diag_1)
```
As we can see, we have many factors and not much we can interpret from them.

First we will need to transform these values from factors to characters (strings), to work more easily
```{r echo=FALSE}
diabetic_data$diag_1 <- as.character(diabetic_data$diag_1)
diabetic_data$diag_2 <- as.character(diabetic_data$diag_2)
diabetic_data$diag_3 <- as.character(diabetic_data$diag_3)
```

Next, we change these values to their group names following the guidelines provided in the Table 2 of the description PDF.

```{r echo=FALSE}
#Circulatory
for (i in 390:459){
  diabetic_data$diag_1[diabetic_data$diag_1 == as.character(i)] <- "Circulatory"
  diabetic_data$diag_2[diabetic_data$diag_2 == as.character(i)] <- "Circulatory"
  diabetic_data$diag_3[diabetic_data$diag_3 == as.character(i)] <- "Circulatory"
}
diabetic_data$diag_1[diabetic_data$diag_1 == "785"] <- "Circulatory"
diabetic_data$diag_2[diabetic_data$diag_2 == "785"] <- "Circulatory"
diabetic_data$diag_3[diabetic_data$diag_3 == "785"] <- "Circulatory"

#Respiratory
for (i in 460:519){
  diabetic_data$diag_1[diabetic_data$diag_1 == as.character(i)] <- "Respiratory"
  diabetic_data$diag_2[diabetic_data$diag_2 == as.character(i)] <- "Respiratory"
  diabetic_data$diag_3[diabetic_data$diag_3 == as.character(i)] <- "Respiratory"
}
diabetic_data$diag_1[diabetic_data$diag_1 == "786"] <- "Respiratory"
diabetic_data$diag_2[diabetic_data$diag_2 == "786"] <- "Respiratory"
diabetic_data$diag_3[diabetic_data$diag_3 == "786"] <- "Respiratory"

#Digestive
name <- "Digestive"
for (i in 520:579){
  diabetic_data$diag_1[diabetic_data$diag_1 == as.character(i)] <- name
  diabetic_data$diag_2[diabetic_data$diag_2 == as.character(i)] <- name
  diabetic_data$diag_3[diabetic_data$diag_3 == as.character(i)] <- name
}
diabetic_data$diag_1[diabetic_data$diag_1 == "787"] <- name
diabetic_data$diag_2[diabetic_data$diag_2 == "787"] <- name
diabetic_data$diag_3[diabetic_data$diag_3 == "787"] <- name


#Injury
name <- "Injury"
for (i in 800:999){
  diabetic_data$diag_1[diabetic_data$diag_1 == as.character(i)] <- name
  diabetic_data$diag_2[diabetic_data$diag_2 == as.character(i)] <- name
  diabetic_data$diag_3[diabetic_data$diag_3 == as.character(i)] <- name
}


#Musculoskeletal
name <- "Musculoskeletal"
for (i in 710:739){
  diabetic_data$diag_1[diabetic_data$diag_1 == as.character(i)] <- name
  diabetic_data$diag_2[diabetic_data$diag_2 == as.character(i)] <- name
  diabetic_data$diag_3[diabetic_data$diag_3 == as.character(i)] <- name
}



#Genitourinary
name <- "Genitourinary"
for (i in 580:629){
  diabetic_data$diag_1[diabetic_data$diag_1 == as.character(i)] <- name
  diabetic_data$diag_2[diabetic_data$diag_2 == as.character(i)] <- name
  diabetic_data$diag_3[diabetic_data$diag_3 == as.character(i)] <- name
}
diabetic_data$diag_1[diabetic_data$diag_1 == "788"] <- name
diabetic_data$diag_2[diabetic_data$diag_2 == "788"] <- name
diabetic_data$diag_3[diabetic_data$diag_3 == "788"] <- name


#Neoplasms
name <- "Neoplasms"
for (i in 140:239){
  diabetic_data$diag_1[diabetic_data$diag_1 == as.character(i)] <- name
  diabetic_data$diag_2[diabetic_data$diag_2 == as.character(i)] <- name
  diabetic_data$diag_3[diabetic_data$diag_3 == as.character(i)] <- name
}


#Diabetes
name <- "Diabetes"
diabetic_data$diag_1[diabetic_data$diag_1 >="250." & diabetic_data$diag_1 <= "250.999999999"] <- name
diabetic_data$diag_2[diabetic_data$diag_2 >="250." & diabetic_data$diag_2 <= "250.999999999"] <- name
diabetic_data$diag_3[diabetic_data$diag_3 >="250." & diabetic_data$diag_3 <= "250.999999999"] <- name
diabetic_data$diag_1[diabetic_data$diag_1 =="250"] <- name
diabetic_data$diag_2[diabetic_data$diag_2 =="250"] <- name
diabetic_data$diag_3[diabetic_data$diag_3 =="250"] <- name


#Other
name <- "Other"
diabetic_data$diag_1[diabetic_data$diag_1 != "Circulatory" & diabetic_data$diag_1 != "Respiratory" & diabetic_data$diag_1 != "Digestive" & diabetic_data$diag_1 != "Diabetes" & diabetic_data$diag_1 != "Injury" & diabetic_data$diag_1 != "Musculoskeletal" & diabetic_data$diag_1 != "Genitourinary" & diabetic_data$diag_1 != "Neoplasms"] <- name
diabetic_data$diag_2[diabetic_data$diag_2 != "Circulatory" & diabetic_data$diag_2 != "Respiratory" & diabetic_data$diag_2 != "Digestive" & diabetic_data$diag_2 != "Diabetes" & diabetic_data$diag_2 != "Injury" & diabetic_data$diag_2 != "Musculoskeletal" & diabetic_data$diag_2 != "Genitourinary" & diabetic_data$diag_2 != "Neoplasms"] <- name
diabetic_data$diag_3[diabetic_data$diag_3 != "Circulatory" & diabetic_data$diag_3 != "Respiratory" & diabetic_data$diag_3 != "Digestive" & diabetic_data$diag_3 != "Diabetes" & diabetic_data$diag_3 != "Injury" & diabetic_data$diag_3 != "Musculoskeletal" & diabetic_data$diag_3 != "Genitourinary" & diabetic_data$diag_3 != "Neoplasms"] <- name


diabetic_data$diag_1 = as.factor(diabetic_data$diag_1)
diabetic_data$diag_2 = as.factor(diabetic_data$diag_2)
diabetic_data$diag_3 = as.factor(diabetic_data$diag_3)
```

The new distribution it follows, as we can see below, is quite more informative:
```{r}
plot(diabetic_data$diag_1)
```

Now the transformation over the three variables is done, we will want to see the summary:

```{r}

print(dim(diabetic_data))
summary(diabetic_data)
```

We will now reduce the names of our variables for better readibility on analysis such as PCA, with the following mapping
```{r echo=FALSE}
keys <- c("admission_type_id", "admission_source_id","discharge_disposition_id",
"time_in_hospital", "medical_specialty", "number_emergency",
"number_inpatient", "num_lab_procedures", "num_procedures",
"num_medications", "number_outpatient","number_diagnoses", "patient_nbr")
shortened <- c("adm_type_id", "adm_source_id","disch_id",
"time_in_hpt", "specialty", "n_emerg",
"n_inp", "n_lab_proc", "n_proc",
"n_med", "n_outp", "n_diag","patient_n")
names(shortened) <- keys
print(data.frame("abbreviated"=shortened))

currCols <- colnames(diabetic_data)
for (currCol in 1:length(currCols)){
  colName <- currCols[currCol]
  if (colName %in% keys){
    colnames(diabetic_data)[currCol] <- shortened[[colName]]
  }
}
```

Now that the deed is done, we take a final look at our data:

```{r}
summary(diabetic_data)
```

And store it already processed, so we don't have to run this code again.

```{r}
write.csv(diabetic_data, file="processed_data.csv")
```